\LARGE \bf \acro: Reducing Supervision in DAgger using SVM-based Risk

### 2015-08-08 paper.tex
added car experiments

\title{\LARGE \bf \acro: Reducing Supervisor Burden using SVM-based Risk to Efficiently Learn Robot Control Policies from Demonstrations }

@@

\begin{abstract}
For robot control problems where no reward function is known and the dynamics of the environment are not specified, the
learning from demonstrations approach provides an algorithmic paradigm for learning a policy from supervisor-provided
demonstrations. Dataset Aggregation (DAgger) in particular yields a state of the art method in this context which treats
the problem of determining a policy as an iterative supervised learning problem which alternates between executing a
trained policy and acquiring new supervisor control inputs for roll-outs of a current best policy estimator. In this
work, we propose a novel algorithm, \acro~which is based on DAgger and which estimates the riskiness of states
encountered by the policy roll-outs by means of an approximate quantile super-levelset estimate of the distribution of
states used for training the current policy. Since learning a quantile function for high-dimensional input data is
challenging, we instead modify an SVN-based approach to quantile estimation that employs regularization to solve this
problem approximately. Our algorithm considers roll-outs with early termination as soon as a state outside a
user-defined quantile super-levelset is encountered. Our approach furthermore reduces the amount of neccessary
supervisor demonstrations by only collecting supervisor control inputs at states outside the current quantile estimate.
In experiments involving a simulated race track and in a Super Mario Atari game benchmark considered by the authors of
DAgger, our approach results in similar performance levels as DAgger while reducing the amount of required supervisor
control inputs by more than {\color{blue}[-- TODO --]} percent.
\end{abstract}

@@

\section{Introduction} 
Consider the problem of teaching a robot a complex task such cooking a meal or as racing a toy car around a track
against human opponents. When the robot is only provided with visual demonstrations without the knowledge or an
underlying cost function describing the task, traditional model based control approaches are often not applicable. In
these situations, the learning from demonstrations (LfD) approach provides a paradigmn for teaching the robot how to
accomplish the task \cite{ross2013learning,pomerleau1989alvinn,schulman2013case}.

@@

The optimization problem in Eq. \ref{eq:LFD_obj} is typically non-convex and difficult to solve since we only have
access to samples from the supervisor's policy $\tilde{\pi}$ arising from the demonstration trajectories $\mathcal{D}$. To address these issues Stephane and Ross proposed an iterative method known as DAgger \cite{ross2010reduction}.

@@

Steps 1 and 2 are repeated for $K$ iterations or until the cumulative surrogate loss approaches zero. Since DAgger
requires rolling out the policy at each iteration, one way to terminate the algorithm is when the performance of the
rolled out policy matches the intended behavior, or more formally when the surrogate loss is below some predefined
threshold $\gamma$. {\color{blue} [-- F: Also talk about $\beta_i$ weights in DAgger --]}.

@@

\subsection{The \acro~algorithm}

@@

 \subsubsection{Step 2}
 In DAgger the policy that was rolled out is $\pi_{\theta_k}$, however this could potential encounter pathological states that are far from the initial set of demonstrations $\mathcal{D}_0$. To prevent this we terminate the rolled out policy if the risk, or distance from the support, is above some threshold $r(\bx_t) > r_0$.  Furthermore, instead of labeling every state now the supervisor is only require to label states outside of the level set or $\bx$, where $g(\bx) = -1$.  These states are then added to the dataset $\mathcal{D}_{k+1}$ and the algorithm goes back to Step 1.

An implicit trade off exists between what how much of the distribution is contained in the selected level set and the amount of exploration the robot is allowed. If the level set is large the robot will be allowed to explore in this region and potentially make more mistakes with respect to the expert, but learn faster. If the level set contains a smaller fraction of data, then the expert will be in control more and the robot will learn at a slower rate.

\section{Theoretical Analysis}
{\color{blue} [-- TEXT FROM HERE IS STILL ROUGH AND CHANGING --]}
\subsection{Convergence Rate}
\todo{These are ideas still in progress}
The convergence rate of DAgger is dependent on the probability of the expert taking over at each iteration $\beta_i$ \cite{ross2010reduction}. For SHEATH, this corresponds to the probability of leaving the support or $p(g(\mathbf{x}) = -1|\pi_{\theta_i})$ . The intuition is that the larger the level set parameter $\nu$ is the lower the probability of leaving is and thus faster convergence rate, but potentially less safe exploration. Not sure if a formal proof is going to be possible though with assuming known dynamics, still looking into it. However, we can experimentally demonstrate this.

\section{Experiments}
\subsection{Driving Example}
A common benchmark in RL is a car driving around a track avoiding other cars \cite{argall2009survey}. In our driving simulator, the car must stay on a rectangular track and dodge the orange cars that are on the track as well. If the car falls off the track, it is placed back on to the center. The car's control space $u = \lbrace -15^\circ, 0, 15^\circ \rbrace$ or 3 discrete commands that change the angle of the car. The internal state space of the car is xy-coordinates and the angle it is facing. Our supervisor is a solver that uses state space search through the driving simulator to plan the next control.

We have the supervisor drive around the track 2 times and then collect the raw images observed. We use a Gaussian Pyramids to down sample the images to $125 \times 125$ RGB pixels and then extract Histogram of Oriented Gradients (HOG) features using OpenCV. For both DAgger and SHEATH, we use a Linear Support Vector Machine (SVM) as the policy representation, with the regularization term on the slack variable $\gamma=0.01$, which was set via cross validation on the initial training examples from our expert. For the optimization of Eq. \ref{eq:dual_sup} in SHEATH we used a radial basis function as well for the kernel with a bandwidth of $\lambda=0.01$ and $\nu = 0.95$ to account for $95\%$ of the training distribution.

We compare performance in terms of minimization of the underlying cost function $c(\bx,\bu)$, which is the sum of the number of times the car left the track and the number of times the car collided with the other cars on the track, versus the number of queries made to the supervisor. Initial results averaged over 1 level shown in Fig.  suggest an $75\%$ reduction in the number of queries needed for SHEATH compared to DAgger.

\begin{figure}[t!]
\centering
\includegraphics[width=6cm, height=6cm]{figures/race_car_track_example.png}
\caption{ An example image from the racing simulator. The car is meant to drive around a rectangular track and dodge other cars. The supervisor drives around the track for 3 iterations and HOG features are extracted from the images.  }

@@

\begin{figure}[t!]
\centering
\includegraphics[width=8cm, height=6cm]{figures/dagger_sheath_no_cars.png}
\caption{ We compare performance in terms of optimization of the underlying cost function $c(\bx,\bu)$, which is the sum of the number of times the car left the track and the number of times the car collided with the other cars on the track. Initial results averaged over 1 level shown suggest an $75\%$ reduction in the number of queries needed for SHEATH compared to DAgger. \todo{Run this over many trials}  }

>>>>

\title{\LARGE \bf \acro: Reducing Supervision in DAgger using SVM-based Risk }

@@

\begin{abstract}
For robot control problems where the reward function  and the dynamics of the environment are not specified, the
learning from demonstrations approach provides an algorithmic paradigm for learning a policy from supervisor-provided
demonstrations. Dataset Aggregation (DAgger) in particular yields a state of the art method in this context which treats
the problem of determining a policy as an iterative supervised learning problem which alternates between executing a
trained policy and acquiring new supervisor control inputs for roll-outs of a current best policy estimator. In this
work, we propose a novel algorithm, \acro~which is based on DAgger and which estimates the riskiness of states
encountered by the policy roll-outs by means of an approximate quantile super-levelset estimate of the distribution of
states used for training the current policy. Since learning a quantile function for high-dimensional input data is
challenging, we instead modify an SVM-based approach to quantile estimation that employs regularization to solve this
problem approximately. Our algorithm considers roll-outs with early termination as soon as a state outside a
user-defined quantile super-levelset is encountered. Our approach furthermore reduces the amount of neccessary
supervisor demonstrations by only collecting supervisor control inputs at states outside the current quantile estimate.
In experiments involving a simulated race track and in a Super Mario Atari game benchmark considered by the authors of
DAgger, our approach results in similar performance levels as DAgger while reducing the amount of required supervisor
control inputs by more than {\color{blue}[-- TODO --]} percent.
\end{abstract}

@@

\section{Introduction} 
Consider the problem of teaching a robot a complex task such cooking a meal or as racing a toy car around a track
against human opponents. When the robot is only provided with visual demonstrations without the knowledge or an
underlying cost function describing the task, traditional model based control approaches are often not applicable. In
these situations, the learning from demonstrations (LfD) approach provides a paradigmn for teaching a robot how to
accomplish the task \cite{ross2013learning,pomerleau1989alvinn,schulman2013case}.

@@

The optimization problem in Eq. \ref{eq:LFD_obj} is  typically non-convex. It can also be difficult to solve, since we only have access to samples from the supervisor's policy $\tilde{\pi}$ arising from the demonstration trajectories $\mathcal{D}$. To address these issues Stephane and Ross proposed an iterative method known as DAgger \cite{ross2010reduction}.

@@

Steps 1 and 2 are repeated for $K$ iterations or until the cumulative surrogate loss approaches zero. Since DAgger
requires rolling out the policy at each iteration, one way to terminate the algorithm is when the performance of the
rolled out policy matches the intended behavior, or more formally when the surrogate loss is below some predefined
threshold $\gamma$ \footnote{In the orginal DAgger the policy rolled out was stochastically mixed with the expert, thus with probability $\beta$ it would either take the expert's action or the robots. The use of this stochasitically mix policy was for theoretical analysis. In practice, it is recommended to set $\beta = 0$ to avoid biasing the sampling~\cite{NIPS2014_5421,ross2010reduction}}

@@

\subsection{SHIV: SVM-Based Reduction in Human Intervention}

@@

 \subsubsection{Step 2}
 In DAgger the policy that was rolled out is $\pi_{\theta_k}$, however this could potential encounter pathological states that are far from the initial set of demonstrations $\mathcal{D}_0$. To prevent this we terminate the rolled out policy if the risk, or distance from the support, is above some threshold $r(\bx_t) > r_0$.  Furthermore, instead of labeling every state now the supervisor is only require to label states outside of the level set or $\bx$, where $g(\bx) = -1$.  These states are then added to the dataset $\mathcal{D}_{k}$

\section{Experiments}
\subsection{Driving Example}
A common benchmark in RL is a car driving around a track avoiding other cars \cite{argall2009survey}. In our driving simulator, the car must stay on the road of polygonal track. If the car falls off the track, it is placed back on to the center. We sample the polygonal track by sampling points from a Gaussian with a mean that is the centered of the video game work space. The convex hull of the sampled points is then computed. The car's control space $u = \lbrace -15^\circ, 0, 15^\circ \rbrace$ or 3 discrete commands that change the angle of the car. The internal state space of the car is xy-coordinates and the angle it is facing. Our supervisor is a solver that uses state space search through the driving simulator to plan the next control.

We have the supervisor drive around the track 2 times and then collect the raw images observed. We use a Gaussian Pyramids to down sample the images to $125 \times 125$ RGB pixels and then extract Histogram of Oriented Gradients (HOG) features using OpenCV. For both DAgger and SHIV, we use a Linear Support Vector Machine (SVM) as the policy representation, with the regularization term on the slack variable $\gamma=0.01$, which was set via cross validation on the initial training examples from our expert. For the optimization of Eq. \ref{eq:dual_sup} in SHIV we used a radial basis function as well for the kernel with a bandwidth of $\lambda=0.01$ and $\nu = 0.95$ to account for $95\%$ of the training distribution.

We compare performance in terms of minimization of the underlying cost function $c(\bx,\bu)$, which is the  number of times the car left the track versus the number of queries made to the supervisor. In Fig. \ref{fig:car_cost}, we plot the performance of DAgger, SHIV and SHIV only using only the traditional One-Class SVM without the modification.  Initial results, which are run for 6 iterations each and are averaged over 40 levels, shown in Fig. \ref{fig:car_cost} suggest an $70\%$ reduction in the number of queries needed for SHIV compared to DAgger.

\begin{figure}[t!]
\centering
\includegraphics[width=6cm, height=6cm]{figures/race_car_track_example.png}
\caption{ An example image from the racing simulator. The car is meant to drive around a rectangular track and dodge other cars. The supervisor drives around the track for 3 iterations and HOG features are extracted from the images.}

@@

\begin{figure}[t!]
\centering
\includegraphics[width=8cm, height=6cm]{figures/dagger_shiv_one_class.eps}
\caption{ We compare performance in terms of optimization of the underlying cost function $c(\bx,\bu)$, which is the number of times the car left the track. Initial results averaged over 30$\%$ level shown suggest an $70\%$ reduction in the number of queries needed for SHEATH compared to DAgger. }
